\begin{thebibliography}{99}

\bibitem{mwiti}{MWITI, Derrick. A 2019 Guide to Semantic Segmentation. In: \textit{Heartbeat} [online]. Fritz AI, 2019 [cit. 2020-06-07]. Dostupné z: \url{https://heartbeat.fritz.ai/a-2019-guide-to-semantic-segmentation-ca8242f5a7fc}}

\bibitem{sergios}{KARAGIANNAKOS, Sergios. Semantic Segmentation in the era of Neural Networks. In: \textit{AI SUMMER} [online]. 2019 [cit. 2020-06-07]. Dostupné z: \url{https://theaisummer.com/Semantic_Segmentation/}}

\bibitem{mehlig}{MEHLIG, Bernhard. Artificial Neural Networks. \textit{ArXiv.org} [online]. 2019, 206 p. [cit. 2020-06-07]. DOI: arXiv:1901.05639. Dostupné z: \url{https://arxiv.org/abs/1901.05639}}

\bibitem{santiago}{PUENTE, Santiago. \textit{Single and Multi-Label Environmental Sound Classification Using Convolutional Neural Networks} [online]. Gothenburg, 2018 [cit. 2020-06-07]. Dostupné z: \url{https://odr.chalmers.se/handle/20.500.12380/255604. Master's thesis. Chalmers University of Technology.}}

%\bibitem{goodfellow}{GOODFELLOW, Ian, Yoshua BENGIO a Aaron COURVILLE. \textit{Deep learning}. Cambridge, Massachusetts: The MIT Press, 2016. ISBN 978-026-2035-613.}

\bibitem{groman}{GROMAN, Martin. \textit{Tvorba umělé neuronové sítě pro výpočet termodynamických veličin} [online]. Brno, 2019 [cit. 2020-06-07]. Dostupné z: \url{http://hdl.handle.net/11012/175381}. Diplomová práce. Vysoké učení technické v Brně. Fakulta strojního inženýrství. Ústav matematiky. Vedoucí práce Tomáš Mauder.}

\bibitem{stanford-github}{CS231n: Convolutional Neural Networks for Visual Recognition: Lecture Notes. \textit{CS231n: Convolutional Neural Networks for Visual Recognition} [online]. Stanford: Stanford University [cit. 2020-06-07]. Dostupné z: \url{https://cs231n.github.io/}}

\bibitem{stanford-L4}{Lecture 4 | Introduction to Neural Networks \textit{YouTube} [online]. 11. August 2018 [cit. 2020-06-07]. Dostupné z: \url{https://www.youtube.com/watch?v=d14TUNcbn1k&list=PL3FW7Lu3i5JvHM8ljYj-zLfQRF3EO8sYv&index=4} }

\bibitem{stanford-L6}{Lecture 6 | Training Neural Networks I \textit{YouTube} [online]. 11. August 2018 [cit. 2020-06-07]. Dostupné z: \url{https://www.youtube.com/watch?v=wEoyxE0GP2M&list=PL3FW7Lu3i5JvHM8ljYj-zLfQRF3EO8sYv&index=6} }

\bibitem{stanford-L7}{Lecture 7 | Training Neural Networks II \textit{YouTube} [online]. 11. August 2018 [cit. 2020-06-07]. Dostupné z: \url{https://www.youtube.com/watch?v=_JB0AO7QxSA&list=PL3FW7Lu3i5JvHM8ljYj-zLfQRF3EO8sYv&index=7} }

\bibitem{stanford-L8}{Lecture 8 | Deep Learning Software \textit{YouTube} [online]. 11. August 2018 [cit. 2020-06-07]. Dostupné z: \url{https://www.youtube.com/watch?v=6SlgtELqOWc&list=PL3FW7Lu3i5JvHM8ljYj-zLfQRF3EO8sYv&index=8} }

\bibitem{stanford-L11}{Lecture 11 | Detection and Segmentation \textit{YouTube} [online]. 11. August 2018 [cit. 2020-06-07]. Dostupné z: \url{https://www.youtube.com/watch?v=nDPWywWRIRo&list=PL3FW7Lu3i5JvHM8ljYj-zLfQRF3EO8sYv&index=11} }

\bibitem{coors}{COORS, Benjamin. \textit{Navigation of Mobile Robots in Human Environments with Deep Reinforcement Learning} [online]. Stockholm, 2016 [cit. 2020-06-08]. Dostupné z: \url{http://www.diva-portal.org/smash/record.jsf?pid=diva2%3A967644&dswid=9005. Degree project. KTH Royal Institute of Technology.}}
		
\bibitem{eniola}{ALESE, Eniola. The curious case of the vanishing \& exploding gradient. In: \textit{Medium} [online]. 2018 [cit. 2020-06-09]. Dostupné z: \url{https://medium.com/learn-love-ai/the-curious-case-of-the-vanishing-exploding-gradient-bf58ec6822eb}}

\bibitem{bushaev}{BUSHAEV, Vitaly. Adam — latest trends in deep learning optimization. In: \textit{Towards Data Science} [online]. Towards Data Science, 2018 [cit. 2020-06-09]. Dostupné z: \url{https://towardsdatascience.com/adam-latest-trends-in-deep-learning-optimization-6be9a291375c}}

\bibitem{issue}{Batch Normalization Issue in SegNet. In: \textit{GitHub} [online]. GitHub, 2017 [cit. 2020-06-09]. Dostupné z: \url{https://github.com/alexgkendall/caffe-segnet/issues/109}}

\bibitem{arvi}{JONNARTH, Arvi. Camera-Based Friction Estimation with Deep Convolutional Neural Networks [online]. Uppsala, 2018 [cit. 2020-06-09]. Dostupné z: \url{https://pdfs.semanticscholar.org/4c35/becacb2aab803468eb38f19d8418d79c7c08.pdf}. Master's thesis. Uppsala Universitet.}

\bibitem{krizhevsky}{KRIZHEVSKY, Alex, Ilya SUTSKEVER a Geoffrey E. HINTON. ImageNet Classification with Deep Convolutional Neural Networks. \textit{Communications of the ACM} [online]. 2017, 60(6), 84-90 [cit. 2020-06-24]. DOI: 10.1145/3065386. ISSN 00010782. Dostupné z: \url{http://search.ebscohost.com.ezproxy.lib.vutbr.cz/login.aspx?direct=true&db=bth&AN=123446102&lang=cs&site=ehost-live}}
		
\bibitem{lecun}{LECUN, Y., L. BOTTOU, Y. BENGIO a P. HAFFNER. Gradient-based learning applied to document recognition. \textit{Proceedings of the IEEE} [online]. 86(11), 2278-2324 [cit. 2020-06-24]. DOI: 10.1109/5.726791. ISSN 00189219. Dostupné z: \url{http://ieeexplore.ieee.org/document/726791/}}

\bibitem{szegedy}{SZEGEDY, Christian, WEI LIU, YANGQING JIA, et al. Going deeper with convolutions. \textit{2015 IEEE Conference on Computer Vision and Pattern Recognition (CVPR)} [online]. IEEE, 2015, 2015, , 1-9 [cit. 2020-06-24]. DOI: 10.1109/CVPR.2015.7298594. ISBN 978-1-4673-6964-0. Dostupné z: \url{http://ieeexplore.ieee.org/document/7298594/}}

\bibitem{vgg}{SIMONYAN, Karen a Andrew ZISSERMAN. Very Deep Convolutional Networks for Large-Scale Visual Recognition. \textit{Visual Geometry Group} [online]. Oxford: University of Oxford, 2014 [cit. 2020-06-11]. Dostupné z: \url{http://www.robots.ox.ac.uk/~vgg/research/very_deep/}}

\bibitem{resnet}{HE, Kaiming, Xiangyu ZHANG, Shaoqing REN a Jian SUN. Deep Residual Learning for Image Recognition. \textit{2016 IEEE Conference on Computer Vision and Pattern Recognition (CVPR)} [online]. IEEE, 2016, 2016, , 770-778 [cit. 2020-06-24]. DOI: 10.1109/CVPR.2016.90. ISBN 978-1-4673-8851-1. Dostupné z: \url{http://ieeexplore.ieee.org/document/7780459/}}

\bibitem{coufal}{COUFAL, J. \textit{Detekce cesty pro mobilní robot analýzou obrazu}. Brno: Vysoké učení
	technické v Brně, Fakulta strojního inženýrství, 2010. 49 s. Diplomová práce. Vedoucí práce: Ing.
	Jiří Krejsa, Ph.D}

\bibitem{bayesian}{KENDALL, Alex, Vijay BADRINARAYANAN a Roberto CIPOLLA. Bayesian SegNet: Model Uncertainty in Deep Convolutional Encoder-Decoder Architectures for Scene Understanding. \textit{ArXiv.org} [online]. 2015, 11 p. [cit. 2020-06-24]. DOI: arXiv:1511.02680. Dostupné z: \url{https://arxiv.org/abs/1511.02680}}

\bibitem{segnet}{KENDALL, Alex, Vijay BADRINARAYANAN a Roberto CIPOLLA. SegNet: A Deep Convolutional Encoder-Decoder Architecture for Image Segmentation. \textit{ArXiv.org} [online]. 2015, 14 p. [cit. 2020-06-24]. DOI: arXiv:1511.00561. Dostupné z: \url{https://arxiv.org/abs/1511.00561}}

\bibitem{segnet_tut}{KENDALL, Alex, Vijay BADRINARAYANAN a Roberto CIPOLLA. SegNet. \textit{Machine Intelligence Laboratory} [online]. Cambridge: University of Cambridge, 2015 [cit. 2020-06-24]. Dostupné z: \url{https://mi.eng.cam.ac.uk/projects/segnet/}}

\bibitem{zeltner}{ZELTNER, Felix. Autonomous Terrain Classification Through Unsupervised Learning [online]. Luleå, 2016 [cit. 2020-06-11]. Dostupné z: \url{http://ltu.diva-portal.org/smash/record.jsf?pid=diva2%3A1051763&dswid=-6301}. Degree project. Luleå University of Technology.}
	
\bibitem{segnet_get_started}{KENDALL, Alex. Getting Started with SegNet. \textit{Machine Intelligence Laboratory} [online]. Cambridge: University of Cambridge, 2015 [cit. 2020-06-11]. Dostupné z: \url{http://mi.eng.cam.ac.uk/projects/segnet/tutorial.html}}

\bibitem{iou}{ROSEBROCK, Adrian. Intersection over Union (IoU) for object detection. In: \textit{Pyimagesearch} [online]. pyimagesearch, 2016 [cit. 2020-06-12]. Dostupné z: \url{https://www.pyimagesearch.com/2016/11/07/intersection-over-union-iou-for-object-detection/}}

\bibitem{filip_github}{SegNet-Tutorial. 2020 [cit. 2020-06-12]. \textit{GitHub} [online]. GitHub. Dostupné z: \url{https://github.com/filipovyfusky/SegNet-Tutorial}}

\bibitem{nvidia}{\textit{NVIDIA} [online]. 2020 [cit. 2020-06-12]. USA: NVIDIA. Dostupné z: \url{https://www.nvidia.com/}}

\bibitem{nvidia_dev}{\textit{NVIDIA Developer} [online]. 2020 [cit. 2020-06-12]. USA: NVIDIA. Dostupné z: \url{https://developer.nvidia.com/}}

\bibitem{caffe}{\textit{Caffe} [online]. Berkeley: Berkeley AI Reseach [cit. 2020-06-12]. Dostupné z: \url{https://caffe.berkeleyvision.org/}}

\bibitem{filip_github_caffe}{caffe-segnet-cudnn5. 2020 [cit. 2020-06-12]. \textit{GitHub} [online]. GitHub. Dostupné z: \url{https://github.com/filipovyfusky/caffe-segnet-cudnn5}}

\bibitem{aizawan_github}{SegNet implementation in Tensorflow. 2020 [cit. 2020-06-12]. \textit{GitHub} [online]. GitHub. Dostupné z: \url{https://github.com/aizawan/segnet}}

\bibitem{labelbox}{\textit{Labelbox} [online]. 2020 [cit. 2020-06-12]. Dostupné z: \url{https://labelbox.com}}

\bibitem{theano}{Convolution arithmetic tutorial. 2018 [cit. 2020-06-12]. \textit{Deep Learning} [online]. LISA lab. Dostupné z: \url{http://deeplearning.net/software/theano_versions/dev/tutorial/conv_arithmetic.html}}

\bibitem{gigabyte}{\textit{GIGABYTE} [online]. 2020 [cit. 2020-06-12]. GIGABYTE. Dostupné z: \url{https://www.gigabyte.com/}}

\bibitem{notes}{MEHLIG, Bernhard. 2019 [cit. 2020-06-12]. FFR135 - \textit{Artificiella neurala nätverk: Lecture Notes}. Gothenburg.}

\end{thebibliography}
